Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-03-20 resex.tex
Removed appendix on 10th-turn FMA. Minor changes to figures.

\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
prstab,
%prstper,
%floatfix,
longbibliography
]{revtex4-1}

@@

% widths for figures
\newlength{\halfwidth}
\setlength{\halfwidth}{0.48\textwidth}
\newlength{\thirdwidth}
\setlength{\thirdwidth}{0.33\textwidth}
\newlength{\bsrtwidth}
\setlength{\bsrtwidth}{0.4\textwidth}
\newlength{\fmawidth}
\setlength{\fmawidth}{0.4\textwidth}

@@

\begin{figure*}
  \begin{minipage}[c]{\columnwidth}
    \includegraphics[width=\columnwidth]{passive_halo_control.png}
    \includegraphics[width=\columnwidth]{active_halo_control.png}
  \end{minipage}
  \hfill
  \begin{minipage}[c]{\columnwidth}
  \includegraphics[width=\columnwidth]{diffusion_illustration}
  \end{minipage}
  \caption{Left: Sketch of passive halo control with a conventional
    collimation system (top) and active halo control, with the
    addition of a hollow electron lens (bottom). Right: Illustration
    of a simplified model of active diffusion enhancement in the
    transverse plane. The diffusion coefficient as a function of
    amplitude (orange) is enhanced in a specific amplitude region when
    the hollow beam is turned on (from solid to dashed line). A
    corresponding reduction in beam tail population (black) is created
    (from solid to dashed line).}
  \label{fig:active_halo_control} 
\end{figure*}

@@

\begin{figure*}
  \includegraphics[width=0.75\textwidth]{hel_layout_epbeam}
  \caption{Layout of the hollow electron lens for HL-LHC. (Courtesy of
    CERN EN-MME mechanical engineering group.)}
  \label{fig:hel_layout}
\end{figure*}

@@

\begin{figure*}
  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} \hfill
  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}
  \caption{Example of current-density distribution measurements for
    the hollow electron gun prototype CHG1b, taken at the Fermilab
    electron lens test stand in
    2017~\cite{hel_res_field_stancari_2017}: 2-dimensional transverse
    profile measurement (left) and calculated 1-dimensional radial
    projection (right).}
  \label{core:fig:0}
\end{figure*}

\begin{figure*}
  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} \hfill
  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
  \caption{Calculated relative electric field for the hollow electron
    gun CHG1b in the transverse $x$-$y$ plane (left) and as
    1-dimensional cuts through the $x$ and $y$ axes (right). The field
    calculations are based on measurements at the Fermilab
    electron-lens test stand combined with Warp calculations of the
    electric potentials and fields in a cylindrical beam pipe.}
\label{core:fig:1}
\end{figure*}

@@

\begin{figure*}
  \begin{tabular}{c}
    \multicolumn{1}{l}{\emph{2016 experiment}} \\
    \includegraphics[width=0.95\textwidth]{bunchfilling_2016.png} \\
    \\
    \multicolumn{1}{l}{\emph{2017 experiment}} \\
    \includegraphics[width=0.95\textwidth]{bunchfilling_2017.png}
  \end{tabular}
  \caption{Bunch filling scheme and excitation patterns for the
    2016~(top) and 2017~(bottom) LHC experiments. In 2016, a total of
    48~bunches was used, whereas in 2017 there were 216~bunches. Each
    bunch is represented by a vertical cyan bar. The bunches were
    grouped in subsets of~4 in 2016 and in subsets of~6 in 2017. Each
    subset experienced the same excitation pattern and amplitude. The
    excitation amplitudes and relative phases are shown in blue or
    red. In 2016, the excitation was only applied in one plane. In
    2017, more injected bunches were allowed without compromising
    machine protection, so it was possible to test all~3 excitation
    planes in the same fill. The transverse damper was active on half
    of the bunches, indicated by the black lines.}
  \label{fig:fill}
\end{figure*}

\begin{figure*}
  \includegraphics[width=\textwidth]{bunchfilling_measured.png}	
  \caption{Example of bunch-by-bunch beam centroid motion for Beam~1
    in the horizontal (B1H, left) and vertical (B1V, right) planes,
    detected by the real-time transverse activity monitor during the
    2017 experiment, when the beam was excited every
    $8^\mathrm{th}$~turn.}
  \label{fig:fill_meas} 
\end{figure*}

@@

\begin{figure*}
  \begin{tabular}{cc}
    no excitation & $10^{\mathrm{th}}$, H+V \\
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
    \\
    7th, H & 7th, V \\
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
  \end{tabular}
  \caption{Frequency-map analysis in betatron tune space based on the
    2016 injection optics with no machine errors and tunes (64.28,
    59.31): without excitation (top left); $10^{\mathrm{th}}$~turn
    pulsing in both horizontal and vertical planes (top left);
    \seventhtp\ in the horizontal (bottom left) and
    vertical (bottom right) planes. The excitation amplitude is
    120~nrad in the corresponding plane. The colors (blue to red)
    represent the tune jitter of tracked particles starting at each
    given location in tune space~\cite{fmalaskar}. The absence of a
    strong excitation of any resonance in case of vertical
    \seventhtp\ and the strong excitation in case
    of horizontal pulsing confirms the excitation of the $7 Q_x$
    resonance. For $10^{\mathrm{th}}$~turn pulsing, there is in
    contrast no significant difference between H, V, or H+V (see
    Appendix~\ref{app:sec:10}, Fig.~\ref{app:fig:fma:10}).}
  \label{fig:patternfma}
\end{figure*}

\begin{figure*}
  \begin{tabular}{cc}
    no excitation & random, H+V \\
    \includegraphics[width=\halfwidth]{2016injerra2b2u_3_5um_hist_y.png} &
    \includegraphics[width=\halfwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
    7th, H+V & $10^{\mathrm{th}}$, H+V \\
    \includegraphics[width=\halfwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
                  &
    \includegraphics[width=\halfwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
  \end{tabular}
  \caption{Calculated beam distributions as a function of vertical
    position from distribution-tracking simulations based on the 2016
    injection optics with $(Q_x, Q_y) = (64.28, 59.31)$ and standard
    lattice errors: no excitation (top left), random excitation (top
    right), \seventhtp\ (bottom left), and
    \tenthtp\ (bottom right). The excitations
    are applied in both planes with an amplitude of 96~nrad. For each
    of the 4~cases, 3~plots are shown. The top plot shows the
    normalized transverse distributions: `initial' (after
    $10^2$~turns, in gray), `final' (after $10^4$~turns, in black),
    and their Gaussian fits (light and dark red, respectively). The
    middle plots show the relative residuals (i.e., differences, in
    percent of the peak value) between final and initial distributions
    (in black) and between each distribution and its Gaussian fit (in
    light and dark red). The ratios between final and initial
    distributions are drawn in black in the bottom plots. The
    residuals emphasize changes near the core of the distributions,
    whereas ratios (when statistically significant), reveal variations
    in the tails.}
  \label{fig:patternhist}
\end{figure*}

@@

\appendix
\section{Additional plots for 10th turn pulsing during the 2016 experiments}
\label{app:sec:10}
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$10^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$10^{\mathrm{th}}$ H
		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skh_dp0_ord10.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$10^{\mathrm{th}}$ V
		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skv_dp0_ord10.png}
	\end{minipage}
	\caption{\label{app:fig:fma:10} FMA analysis for \tenthtp\ based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31). The excitation is 120~nrad in the corresponding plane. There is no significant difference in the FMA analysis between pulsing only in H, only in V or in H+V.}
\end{figure}

\bibliography{resex-bibdesk}

>>>>

\documentclass[%
%preprint,
reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
prstab,
%prstper,
floatfix,
%endfloats,
longbibliography
]{revtex4-1}

@@

% widths for figures
\newlength{\halfwidth}
\setlength{\halfwidth}{0.47\textwidth}
\newlength{\thirdwidth}
\setlength{\thirdwidth}{0.33\textwidth}
\newlength{\bsrtwidth}
\setlength{\bsrtwidth}{0.38\textwidth}
\newlength{\fmawidth}
\setlength{\fmawidth}{0.38\textwidth}
\newlength{\smallfmawidth}
\setlength{\smallfmawidth}{0.24\textwidth}

@@

\begin{figure*}
  \hspace*{\fill}
  \begin{minipage}[c]{0.45\textwidth}
    \includegraphics[width=\linewidth]{passive_halo_control.png}
    \includegraphics[width=\linewidth]{active_halo_control.png}
  \end{minipage} \hfill
  \begin{minipage}[c]{0.4\textwidth}
    \includegraphics[width=\linewidth]{diffusion_illustration}
  \end{minipage}
  \hspace*{\fill}
  \caption{Left: Sketch of passive halo control with a conventional
    collimation system (top) and active halo control, with the
    addition of a hollow electron lens (bottom). Right: Illustration
    of a simplified model of active diffusion enhancement in the
    transverse plane. The diffusion coefficient as a function of
    amplitude (orange) is enhanced in a specific amplitude region when
    the hollow beam is turned on (from solid to dashed line). A
    corresponding reduction in beam tail population (black) is created
    (from solid to dashed line).}
  \label{fig:active_halo_control} 
\end{figure*}

@@

\begin{figure*}
  \includegraphics[width=0.8\textwidth]{hel_layout_epbeam}
  \caption{Layout of the hollow electron lens for HL-LHC. (Courtesy of
    CERN EN-MME mechanical engineering group.)}
  \label{fig:hel_layout}
\end{figure*}

@@

\begin{figure*}
  \begin{tabular}{cc}
    \includegraphics[width=0.48\textwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} &
    \includegraphics[width=0.4\textwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}\\
  \end{tabular}
  \caption{Example of current-density distribution measurements for
    the hollow electron gun prototype CHG1b, taken at the Fermilab
    electron lens test stand in
    2017~\cite{hel_res_field_stancari_2017}: 2-dimensional transverse
    profile measurement (left) and calculated 1-dimensional radial
    projection (right).}
  \label{core:fig:0}
\end{figure*}

\begin{figure*}
  \begin{tabular}{cc}
    \includegraphics[width=0.4\textwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} &
    \includegraphics[width=0.4\textwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}\\
  \end{tabular}
  \caption{Calculated relative electric field for the hollow electron
    gun CHG1b in the transverse $x$-$y$ plane (left) and as
    1-dimensional cuts through the $x$ and $y$ axes (right). The field
    calculations are based on measurements at the Fermilab
    electron-lens test stand combined with Warp calculations of the
    electric potentials and fields in a cylindrical beam pipe.}
  \label{core:fig:1}
\end{figure*}

@@

\begin{figure*}
  \begin{tabular}{c}
    \multicolumn{1}{l}{\emph{2016 experiment}} \\
    \includegraphics[width=0.9\textwidth]{bunchfilling_2016.png} \\
    \\
    \multicolumn{1}{l}{\emph{2017 experiment}} \\
    \includegraphics[width=0.9\textwidth]{bunchfilling_2017.png}
  \end{tabular}
  \caption{Bunch filling scheme and excitation patterns for the
    2016~(top) and 2017~(bottom) LHC experiments. In 2016, a total of
    48~bunches was used, whereas in 2017 there were 216~bunches. Each
    bunch is represented by a vertical cyan bar. The bunches were
    grouped in subsets of~4 in 2016 and in subsets of~6 in 2017. Each
    subset experienced the same excitation pattern and amplitude. The
    excitation amplitudes and relative phases are shown in blue or
    red. In 2016, the excitation was only applied in one plane. In
    2017, more injected bunches were allowed without compromising
    machine protection, so it was possible to test all~3 excitation
    planes in the same fill. The transverse damper was active on half
    of the bunches, indicated by the black lines.}
  \label{fig:fill}
\end{figure*}

\begin{figure*}
  \includegraphics[width=0.9\textwidth]{bunchfilling_measured.png}	
  \caption{Example of bunch-by-bunch beam centroid motion for Beam~1
    in the horizontal (B1H, left) and vertical (B1V, right) planes,
    detected by the real-time transverse activity monitor during the
    2017 experiment, when the beam was excited every
    $8^\mathrm{th}$~turn.}
  \label{fig:fill_meas} 
\end{figure*}

@@

\begin{figure*}
  \begin{tabular}{cc}
    no excitation & $10^{\mathrm{th}}$, H+V \\
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
    \\
    7th, H & 7th, V \\
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
  \end{tabular}
  \caption{Frequency-map analysis in betatron tune space based on the
    2016 injection optics with no machine errors and tunes (64.28,
    59.31): without excitation (top left); $10^{\mathrm{th}}$~turn
    pulsing in both horizontal and vertical planes (top left);
    \seventhtp\ in the horizontal (bottom left) and vertical (bottom
    right) planes. The excitation amplitude is 120~nrad in the
    corresponding plane. The colors (blue to red) represent the tune
    jitter of tracked particles starting at each given location in
    tune space~\cite{fmalaskar}. The absence of a strong excitation of
    any resonance in case of vertical \seventhtp\ and the strong
    excitation in case of horizontal pulsing confirms the excitation
    of the $7 Q_x$ resonance. For $10^{\mathrm{th}}$~turn pulsing,
    there is in contrast no significant difference between H, V, or
    H+V (Fig.~\ref{fig:fma:10}).}
  \label{fig:patternfma}
\end{figure*}

\begin{figure}
  \begin{tabular}{cc}
    No excitation & 10th, H+V \\
    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png} \\
    10th, H & 10th, V \\
    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerrut10skh_dp0_ord10.png} &
    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerrut10skv_dp0_ord10.png} \\
  \end{tabular}
  \caption{FMA for \tenthtp\ based on the 2016 injection
    optics with no lattice errors and tunes (64.28, 59.31). The
    excitation is 120~nrad in the corresponding plane. There is no
    significant difference between pulsing in H only, V only, or in H+V.}
  \label{fig:fma:10}
\end{figure}

\begin{figure*}
  \begin{tabular}{cc}
    No excitation & Random, H+V \\
    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_3_5um_hist_y.png} &
    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
    7th, H+V & 10th, H+V \\
    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
                  &
    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
  \end{tabular}
  \caption{Calculated beam distributions as a function of vertical
    position from distribution-tracking simulations based on the 2016
    injection optics with $(Q_x, Q_y) = (64.28, 59.31)$ and standard
    lattice errors: no excitation (top left), random excitation (top
    right), \seventhtp\ (bottom left), and
    \tenthtp\ (bottom right). The excitations
    are applied in both planes with an amplitude of 96~nrad. For each
    of the 4~cases, 3~plots are shown. The top plot shows the
    normalized transverse distributions: `initial' (after
    $10^2$~turns, in gray), `final' (after $10^4$~turns, in black),
    and their Gaussian fits (light and dark red, respectively). The
    middle plots show the relative residuals (i.e., differences, in
    percent of the peak value) between final and initial distributions
    (in black) and between each distribution and its Gaussian fit (in
    light and dark red). The ratios between final and initial
    distributions are drawn in black in the bottom plots. The
    residuals emphasize changes near the core of the distributions,
    whereas ratios (when statistically significant), reveal variations
    in the tails.}
  \label{fig:patternhist}
\end{figure*}

@@

% bibliography
\bibliography{resex-bibdesk}
